Literature Review

### 2019-02-21 background.Rmd
update

Undergraduate introductory STEM courses typically include technical writing training. Often it is treated as an incidental skill that each student develops "along the way." Or, the instructional strategy is  haphazardly designed, implemented, and executed.

Adding to the challenge, faculty may not be directly responsible for helping students develop writing skills. Technical writing training usually is part of a lab course associated with a lecture class. Many times lab sections are led by graduate teaching assistants or even advanced undergraduates. TAs may have under-developed writing skills themselves, lack pedagogical content knowledge needed to guide students effectively, or not have been explicitly trained to teach writing in a particular way.

@@

## Context for This Project
From prior studies (REFs), we know students make greater gains as writers if instructors avoid telling students what to correct ("copy editing"), and use a coaching-oriented approach instead. To that end, we train graduate teaching assistants to use a **global issues first, specific issues later** approach to correcting undergraduate technical writing. We ask TAs to spend a limited block of time writing fewer comments that focus on broader issues. Ideally, TAs provide no more than 3-5 comments per text page, and comments should encourage student reflection. As students' writing improves, the TA can focus on smaller issues. Our bins-based report grading protocol reinforces this strategy.

The coaching-oriented approach is well-supported by prior research, but we are uncertain how well TAs can implement the recommended strategy. Effort self-reporting by TAs is likely to be unreliable. We have repeatedly asked ourselves, **"what do TAs ACTUALLY spend most of their time and effort making comments about when they grade student writing?"**

## Prior Work
In 2018, we began trying to answer this question by analyzing TAs' comments from student papers directly. We extracted and tabulated all comments from approximately 500 pairs of reports. After initial data cleaning (removing duplicates, splitting comments addressing multiple independent writing issues, and removing comments lacking any context for judging intent), we had a working dataset of ~11,000 separate comments.

We next used a subset of 100 randomly selected comments to build an initial qualitative codebook for sorting comments into types. The initial codebook was re-tested using 1100 additional randomly selected comments. Using the final codebook, we sorted all 11,000 comments into separate categories (see LINK for complete codebook.)

Coded TA comments were sorted on two axes:

@@

Raw counts for each subject/structure pair were converted to fractional frequencies. The final table (excerpt is shown below) provided us with a direct estimate of TA effort.

@@

## The Central Problem of This Project
Coding TA comments this way is informative but doing so by hand is unsustainable. Developing the codebook then rating the initial set of 11,000 comments required nearly 2 months of investigator effort. Even with the codebook in place, individual coders need training to achieve sufficient accuracy, and there is significant risk of "coding drift" over time.

@@

>>>>

Undergraduate introductory STEM courses often include technical writing training. Too often it is treated as an incidental skill that each student develops "along the way." At other times the instructional strategy is  haphazardly designed, implemented, and executed.

Adding to the learning challenge, faculty may not be the ones helping students develop writing skills. If technical writing training is part of a lab course, graduate teaching assistants or even advanced undergraduates may be who teaches technical writing. TAs may have under-developed writing skills themselves, lack pedagogical content knowledge needed to guide students effectively, or not have been explicitly trained to teach writing in a particular way.

@@

####\  
##Context & Prior Work
From prior studies (REFs), we know students make greater gains as writers if instructors avoid telling students what to correct ("copy editing"), and use a coaching-oriented approach instead. To that end, our program trains graduate TAs to use a **global issues first, specific issues later** approach to correcting undergraduate technical writing. We ask them to spend a shorter amount of time writing fewer comments that focus on broader issues. Ideally, TAs provide no more than 3-5 comments per text page, and the comments given should encourage student reflection. As students' writing improves, TAs can focus on smaller issues. Our bins-based report grading protocol reinforces this strategy.

Our coaching-oriented approach is well-supported by prior research, but we are uncertain how accurately TAs implement the recommended strategy. We have repeatedly asked ourselves, **"what do TAs ACTUALLY spend most of their time and effort making comments about when they grade student writing?"** Self-reported TA effort is likely to be unreliable. What we need is a more direct measure of effort.

In 2018, we began trying to analyze TAs' comments from student papers directly. We extracted and tabulated all comments from approximately 500 pairs of reports. After initial data cleaning (removing duplicates, splitting comments addressing multiple independent writing issues, and removing comments lacking any context for judging intent), we had a working dataset of ~11,000 separate comments.

We next used a subset of 110 randomly selected comments to build an initial qualitative codebook for classifying comments based on subject and structure. This initial codebook was tested and revised using blocks of 1100 additional randomly selected comments until the categories and criteria stabilized. Using the final codebook, we then sorted all 11,000 comments into separate categories (see PAGE_LINK for complete codebook.)

TA comments were coded and sorted according to:

@@

Raw counts for each subject/structure pair were then converted to fractional frequencies. The final frequency table (excerpted below) provided us with a direct estimate of TA effort and emphasis.

@@

##Central Problem
Analyzing TA comments this way is informative but coding by hand is unsustainable. Developing the codebook then rating the initial set of 11,000 comments required nearly 80 hours of investigator effort. Even with the codebook in hand, individual coders need training to achieve sufficient accuracy, and there is significant risk of "coding drift" over time.

@@

##Work Plan
The work plan for this project is:

1.  Compile initial dataset and de-identify both TAs and students by replacing names with anonymous IDs.
2.  Import TA comments and relevant metadata from student reports into R as a "tidy" dataframe.
3.  Explore comment data structure, n-gram frequencies, etc., to identify potentially useful elements for feature engineering.
4.  Write an initial Naive Bayes workflow for supervised text classification.
5.  Assign TA comments to categories using the NB classifier, then compare outcomes from automated classification to hand-coded results.
6.  Using features (from Step #3) and comparison data from initial classification protocol (from Step #4) as guides, test iterations and optimize the classifier method. 
7.  Apply the optimized classifier to the original comment dataset, identify all comments that are being classified incorrectly, and look for potential patterns.
8.  Write a small set of rule- or REGEX pattern-based searches that can identify comments which are more likely to be classified incorrectly. 
9.  Append rules/pattern-based pre-screening to the optimized Naive Bayes classifier.
10.  Validate this vertical Ensemble Classifier by classifying a subset of ~2000 new TA comments extracted from a non-overlapping set of student lab reports. Compare automated results for new data set to hand-coded categories.

### 2019-02-21 findings.Rmd
update

---
title: "Project"
author: "Dan Johnson"
date: "10/26/2018"
output: html_document
bibliography: library.bib
biblio-style: apalike
---
```{r, child="_setup.Rmd"}
```

# Abstract

@@

# Introduction
## Background
Undergraduate introductory STEM courses typically include technical writing training. Often it is treated as an incidental skill that each student develops "along the way." Or, the instructional strategy is  haphazardly designed, implemented, and executed. Adding to the challenge, faculty may not be directly responsible for helping students develop writing skills.

Technical writing training usually is part of a lab course associated with a lecture class. These lab sections may be be led by graduate teaching assistants or even advanced undergraduates. These instructors may have under-developed writing skills themselves, or lack pedagogical content knowledge needed to guide students effectively.

In 2014 we launched an automated data collection process we call SAWHET that collects systematic observations and sample data from student writers and graduate instructors. Our resarch goal is to better understand student and instructor behaviors within the technical writing instructional space. Our practical goal is to develop techniques for teaching technical writing that:

*  Can be implemented at large scale by less experienced teaching assistants;
*  Emphasize student development through holistic coaching;
*  Shift student and instructor attention away from simple copy-correction towards larger writing issues; and
*  Embody other best practices outlined in nearly 40 years of research literature on "writing in the disciplines."

We have a structured database of >5000 laboratory reports written by undergraduate students for introductory biology courses. In addition we have all comments that a TA made on the individual reports, plus associated metadata:

*  Context. Was it first or second report of the semester? Initial submission or revision? What course and topic in the course? What year in college was the author?
*  Instructor information. Which TA provided the feedback? What score was assigned, and what what the rationale behind that score?

We are using multiple text-mining methods to map key features of student writing, as well as connections and correlations between those features. These insights then inform how we teach technical writing and train graduate instructors.

@@

## Context for This Project & Prior Work
Students make greater gains as writers if instructors avoid telling students what to correct ("copy editing"), and use a coaching-oriented approach instead. To that end, we train graduate teaching assistants to use a **global first, specific later** approach to grading undergraduate technical writing. We ask TAs to spend a limited block of time writing fewer comments that focus on broader issues. Ideally, TAs provide no more than 3-5 comments per text page, and comments should encourage student reflection. As students' writing improves, the TA can focus on smaller issues.

This strategy is well-supported by prior research, but we are uncertain how well TAs can implement the recommended strategy. One of our ongoing questions has been, **"what do TAs ACTUALLY spend most of their time and effort making comments about when they grade student writing?"**

In Summer 2018, we began analyzing TAs' comments from student papers. We extracted and tabulated all comments from approximately 500 pairs of reports (initial submissions and revisions) collected during a single semester in 3 different courses. Initial data cleaning involved removing duplicates, splitting comments addressing multiple independent writing issues, and removing comments lacking any context for judging intent. In all, the final tabulated dataset included ~11,000 separate comments.

We next used a subset of 100 randomly selected comments to build an initial qualitative codebook for sorting comments into types. The initial codebook was re-tested using 1100 additional randomly selected comments. Using the final codebook, we sorted all 11,000 comments into separate categories (see LINK for complete codebook.)

Coded TA comments were sorted on two axes:

* __Subject__: did the comment focus mainly on Basic criteria, Technical issues, Writing quality, Logic, or Other issues
* __Structure__: was the comment mainly Copy correction, General or Specific information, Pointing to another resource, or Holistic coaching for improvement?

The sorted contingency table provided us with a relative estimate of TA effort.

@@

## The Central Problem of This Project
Qualitative coding of TA comments is informative but unsustainable. Developing the codebook then rating the first set of 11,000 comments required nearly 2 months of investigator effort. Even with the codebook, individual coders will require training to achieve any accuracy. Given 10-12,000 comments are generate EACH SEMESTER, an automated sorting method is needed.

@@

# Literature Review
##General Approach to Initial Dataset
The following excerpt from *Text Mining with R: A Tidy Approach* (Julia Silge and David Robinson) summarizes how extended text (in this case, TA comments) can be organized as a tidy matrix.*All emphases are mine.*

>We thus define the tidy text format as being a table with one-token-per-row. A token is a **meaningful unit of text**, such as a word, that we are interested in using for analysis, and tokenization is the process of splitting text into tokens.

>This one-token-per-row structure is in contrast to the ways text is often stored in current analyses, perhaps as strings or in a document-term matrix. For tidy text mining, the token that is stored in each row is most often a single word, **but can also be an n-gram, sentence, or paragraph** . In the tidytext package, we provide functionality to tokenize by commonly used units of text like these and convert to a one-term-per-row format.

Using tidy format (vs. other common text data formats) better maintains relationships between individual comments and metadata, which simplifies subsequent analysis.

@@

# Methods

@@

## Data

@@

## Analysis

@@

# Results

# Discussion

# Conclusion

# References

>>>>

---
title: "Findings"
author: "Dan Johnson"
date: "2/22/2018"
output: html_document
bibliography: library.bib
biblio-style: apalike
---
```{r, child="_setup.Rmd"}
```

##Classifier Accuracy
The tables below show observed frequencies for 9,340 TA comments coded by hand, and frequencies for the same TA comments categorized using the optimized Naive Bayes classifier **without** pre-screening. The far right columns are overall subject frequencies, the bottom rows, overall structure frequencies.

@@

![*Comparison of frequency tables generated by hand-coding TA comments (top) versus predicted categories using optimized NB classifier (bottom). *](/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/default_website/images/Result_of_NB_2D_mapping_Feb2019.png)

@@

On average, the classifier is only 80-90% accurate for one subject sub-category or one specific comment. OVERALL though, the optimized NB classifier estimated relative frequencies of Subject subcategories very well; there was <1.5% difference in fractional frequencies between observed (hand-coded) and computationally assigned Subject sub-categories.

Overall accuracy was considerably lower for Structure sub-categories. Specifically, the NB classifier over-assigned comments to the "Copy-editing" sub-category, and under-assigned comments to the "Specifics" sub-category. Most incorrectly classified comments belonged to the "Technical" Subject sub-category. This **suggest** there may be specific terms present in comments within this sub-category that are problematic, and warrant further exploration.

@@

##Data Story 2

@@

#Tool or Analysis 2

@@

##Data Story 3

@@

##Data Story 4

@@

@@

### 2019-02-21 litreview.Rmd
update

>>>>

I found this summary, and a description of the approach at
[https://www.tidytextmining.com/topicmodeling.html]():

>"In text mining, we often have collections of documents, such as blog posts or news articles, that we’d like to divide into natural groups so that we can understand them separately. Topic modeling is a method for unsupervised classification of such documents, similar to clustering on numeric data, which finds natural groups of items even when we’re not sure what we’re looking for.

>"Latent Dirichlet allocation (LDA) is a particularly popular method for fitting a topic model. It treats each document as a mixture of topics, and each topic as a mixture of words. This allows documents to “overlap” each other in terms of content, rather than being separated into discrete groups, in a way that mirrors typical use of natural language."

My challenge is that LDA clusters items based on its own analysis. My groups are already sorted based on the **existing coding book.** The individual groups are already labeled.

####\

**Option: Comparison Modeling Based on Twitter Strategies**

What if I treated each of the pre-defined sub-sets of comments I created using codebook analysis as if they were tweets from separate users? That approach is described here: 
	https://www.tidytextmining.com/twitter.html

Strategy would be:

*  Un-nest tokens to create a tidy frame and remove stop words using anti-join
*  Calculate word frequencies for each sub-type
*  Compare frequencies by calculating log-odds ratios

####\

**Option: Comparison Modeling Via Text Classification**

What if I treated each of the comments as text blocks where pre-defined words are more or less likely to be associated with the category. Jerid developed a Shiny app that does something similar:
https://francojc.github.io/2015/08/15/generating-annotated-text-in-shiny/

####\

**Option: Latent Dirichlet Allocation**

# Session 4 (Dec. 3, 2018)
Looking at Jerid's *Intro to Statistical Thinking* clarified some points in my approach. I definitely need to be thinking about this problem more in terms of unsupervised exploration. It did raise some follow-up questions:

*  A training set with 80% of the data seems excessive. If there is a large dataset, is there any cost or benefit to splitting to 33/33/34%, or even 25% x 4? The 80/20 and 60/20/20% also seems out of balance, but there is probably logic I do not see. 
*  In the exploration phase, is re-sampling such a major concern?

**Separate note:** thank you many times over for the Regex101 link! Writing regex correctly has long been a pain for me! This greatly simplifies the process.

####\
